---
title: "Activity Report"
author: "DaniGCFH"
date: "2024-01-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(dplyr)

```

## What is mean total number of steps taken per day?

For this part of the assignment, I'll first read the data and process it to have a view on the daily steps as shown in the histogram.

```{r daily activity}
data <- read.csv("activity.csv")
data$date <- as.Date(data$date)
days <- unique(data$date)
df <- data.frame(matrix(ncol=2))
colnames(df) <- c("days", "steps")

for (day in days){
  steps<-sum(data$steps[data$date==day])
  if (is.na(df$days[1])==TRUE){
    df$days <- as.Date(day)
    df$steps <- as.numeric(steps)
  }
  else{
    row <- c(as.Date(day), as.numeric(steps))
    df <- rbind(df, row)
  }
}

hist(df$steps, xlab = "Number of steps", main = NULL)
knitr::opts_chunk$set(echo = FALSE)


```

The median number of steps taken is `r median(df$steps, na.rm=TRUE)`and the mean number of steps taken is `r mean(df$steps, na.rm=TRUE)`

## What is the average daily activity pattern?

For this part of the assignment, I repeated the data processing, this time by interval, to produce the required plot.

```{r base activity pattern}
intervals <- unique(data$interval)
di <- data.frame(matrix(ncol=2))
colnames(di) <- c("interval", "steps")

for (interval in intervals){
  steps<-mean(data$steps[data$interval==interval], na.rm=TRUE)
  if (is.na(di$interval[1])==TRUE){
    di$interval <- as.numeric(interval)
    di$steps <- as.numeric(steps)
  }
  else{
    row <- c(as.numeric(interval), as.numeric(steps))
    di <- rbind(di, row)
  }
}

plot(di, type="l", xlab= "Interval", ylab = "Average Number of Steps")
knitr::opts_chunk$set(echo = FALSE)


```

The 5 minute interval with the maximum average number of steps accross all the days in the dataset is the interval number `r di$interval[di$steps==max(di$steps)]`

## Imputing missing values

```{r missing values}

rowNA <- table(!complete.cases(data))
knitr::opts_chunk$set(echo = FALSE)

```

The number of rows with missing values is `r rowNA[2]`

To fill the missing values in the dataset, I replaced all missing values for the average number of steps per interval.

```{r replacing}
data1 <- data
for (i in c(1:17568)){
  if (is.na(data1$steps[i])){
    data1$steps[i] <- di$steps[di$interval==data1$interval[i]]
  }
}

days <- unique(data1$date)
df1 <- data.frame(matrix(ncol=2))
colnames(df1) <- c("days", "steps")

for (day in days){
  steps<-sum(data1$steps[data1$date==day])
  if (is.na(df1$days[1])==TRUE){
    df1$days <- as.Date(day)
    df1$steps <- as.numeric(steps)
  }
  else{
    row <- c(as.Date(day), as.numeric(steps))
    df1 <- rbind(df1, row)
  }
}
hist(df1$steps, xlab = "Number of steps", main = NULL)
knitr::opts_chunk$set(echo = FALSE)


```

The median number of steps taken is `r median(df1$steps, na.rm=TRUE)` and the mean number of steps taken is `r mean(df1$steps, na.rm=TRUE)`

These results are very close from the initial results, but differ in that the median value now is equal to the mean value.

## Are there differences in activity patterns between weekdays and weekends?

For this part of the assignment, I repeated the data processing while subsetting the database between weekdays and weekends.

```{r dif patterns}
data$weekday <- weekdays(data$date)
data$weekend[data$weekday =="Sunday" |data$weekday =="Saturday"] <- TRUE
data$weekend[data$weekday !="Sunday" & data$weekday !="Saturday"] <- FALSE
weekday <- subset(data, weekend==FALSE)
weekend<- subset(data, weekend==TRUE)
intervals <- unique(data$interval)
dend <- data.frame(matrix(ncol=2))
dday <- data.frame(matrix(ncol=2))
colnames(dend) <- c("interval", "steps")
colnames(dday) <- c("interval", "steps")

for (interval in intervals){
  steps<-mean(weekday$steps[weekday$interval==interval], na.rm=TRUE)
  if (is.na(dday$interval[1])==TRUE){
    dday$interval <- as.numeric(interval)
    dday$steps <- as.numeric(steps)
  }
  else{
    row <- c(as.numeric(interval), as.numeric(steps))
    dday <- rbind(dday, row)
  }
}

for (interval in intervals){
  steps<-mean(weekend$steps[weekend$interval==interval], na.rm=TRUE)
  if (is.na(dend$interval[1])==TRUE){
    dend$interval <- as.numeric(interval)
    dend$steps <- as.numeric(steps)
  }
  else{
    row <- c(as.numeric(interval), as.numeric(steps))
    dend <- rbind(dend, row)
  }
}

par(mfcol = c(2,1)) 
plot(dday$interval, dday$steps, type="l", xlab= "Interval", ylab = "Average Number of Steps")
plot(dend$interval, dend$steps, type="l", xlab= "Interval", ylab = "Average Number of Steps")
knitr::opts_chunk$set(echo = FALSE)


```

The plot show that the weekdays have a single peak in steps while the weekends seem to have multiple peaks accros the day
